Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2011-11-29 extended_catalog/paper/extended_catalog.tex
new version
updateing text

\begin{abstract}
Spatial extension is an important characteristic for correctly
associating $\gamma$-ray-emitting sources with their counterparts at other wavelengths
and for obtaining an unbiased model of their spectra.  We present a new
method for quantifying the spatial extension of sources detected by the Large
Area Telescope (LAT), the primary science instrument on the {\em \fermi
Gamma-ray Space Telescope} (\fermi).  We perform a series of Monte Carlo
simulations to validate this tool and calculate the LAT threshold for
detecting the spatial extension of sources.  
\newtext{
We then test all sources
in the second \fermi-LAT catalog (2FGL) for extension. 
We report the detection of seven new spatially extended sources.
}
\end{abstract}

@@

A number of astrophysical source classes including supernova remnants
(SNRs), pulsar wind nebulae (PWNe), molecular clouds, normal galaxies,
and galaxy clusters are expected to be spatially resolvable at
\gev energies.  Additionally, dark matter satellites are also hypothesized to
be spatially extended.  The Large Area Telescope
(LAT) on the {\em \fermi Gamma-ray Space Telescope} (\fermi) has
detected seven SNRs which are significantly extended
at \gev energies: 
W51C, W30, 
IC443, W28, W44, RX\,J1713.7$-$3946,
and the Cygnus Loop
\citep{w51c,w30_lat,ic443,w28,w44,rx_j1713_lat,cygnus_loop_lat}. In addition, 
three extended PWN have been detected: MSH\,15$-$52, Vela~X, and
HESS\,J1825$-$137 \citep{msh1552,velax,fermi_hess_j1825}. Two
nearby galaxies, the Large and Small Magellanic Clouds, and one
radio galaxy, Centaurus A, were spatially resolved at \gev energies
\citep{lmc,smc,cen_a_lat}.  A number of additional sources detected
at \gev energies are positionally coincident with sources that exhibit
large enough extension at other wavelengths to be spatially
resolvable by the LAT at \gev energies.

@@

\newtext{
A new maximum-likelihood analysis tool has been developed to address the
unique requirements for studying spatially extended sources with the LAT.
It works by maximizing the Poisson likelihood to find the observing counts
given a parametrized spatial and spectral model of the sky.  
Space is
divided into cubes of position and energy using the HEALPix representation of
the sky \citep{healpix} and the likelihood is maximized over all bins in
a region.
}
The extension of a source can be modeled by a geometric shape
(e.g. a disk or a two-dimensional Gaussian) and the position, extension,
and spectrum of the source can be simultaneously fit.

This type of analysis is not feasible using the standard LAT likelihood
analysis tool \gtlike\footnote{\gtlike is distributed publicly by the
FSSC.} because it can only fit the spectral parameters of the model
unless a more sophisticated iterative procedure is used.  
We note that \gtlike has been used in the
past in several studies of source extension in the LAT Collaboration
\citep{lmc,smc,w28,w51c}.  In these studies, a profile based upon a
set of \gtlike maximum likelihood fits at fixed extensions was used
to build a profile of the likelihood as a function of extension.
\newtext{
The \gtlike likelihood profile approach has been show to correctly
reproduce the extension of simulated extended sources assuming that the
true position is known \citep{francesco_2011}.  But it is not optimal
because the position, extension, and spectrum of the source must be
simultaneously fit to find the best fit parameters and to maximize the
statistical significance of the detection.  Furthermore, because the \gtlike
approach is computationally intensive, no large-scale Monte Carlo
simulations have been run to calculate its false detection rate.
}

@@

In \pointlike, it is assumed that the spatial and spectral model
of an extended source are separable, i.e. that the source model
$M(l,b,E)=S(l,b)\times X(E)$ where $S(l,b)$ is the spatial distribution
and $X(E)$ is the spectral distribution.  To fit an extended source,
\pointlike convolves the extended source shape with the PSF (as a function
of energy) and uses the \minuit library \citep{minuit_documentation}
to maximize the likelihood by simultaneously varying the position,
extension, and spectrum of the source.  As will be described in
Section~\ref{monte_carlo_validation}, simultaneously fitting the
position, extension, and spectrum is important to maximize
the statistical significance of the detection of the extension of a source.
To avoid projection effects, the longitude and latitude 
of the source are
not directly fit but instead the displacement of the source in a rotated
reference frame.

@@

For extended sources with an assumed radially-symmetric shape,
we optimized the calculation by performing one
of the integrals analytically.
The expected photon 
distribution can be written as
\begin{equation}
  \text{PDF}(\vec r) = \int  \text{PSF}(|\vec r - \vec r'|)I_\text{src}(\vec r') r' dr' d\phi'
\end{equation}
where $I_\text{src}(\vec r')$ is the spatial distribution of the
source.
\newtext{
The PSF of the LAT is currently parameterized 
in the Pass~7\_V6 (P7\_V6) Source Instrument
Response Function \citep[IRFs,][]{lat_on_orbit_psf} by a King function \citep{king_function}:
}
\begin{equation}
  \text{PSF}(r) = 
  \frac{1}{2\pi\sigma^2}
  \left(1-\frac{1}{\gamma}\right)
  \left(1+\frac{u}{\gamma}\right)^{-\gamma},
\end{equation}
where $u=(r/\sigma)^2/2$ and $\sigma$ and $\gamma$ are free parameters
\citep{matthew_kerr_thesis}.  For radially-symmetric extended sources,
the angular part of the integral can be evaluated analytically
\begin{align}
  \text{PDF}(u) & = \int_0^\infty r' dr'
  I_\text{src}(v) 
  \int_0^{2\pi} d\phi' 
  \text{PSF}(\sqrt{2\sigma^2(u+v-2\sqrt{uv}\cos(\phi-\phi'))})
  \\
  & = \int_0^\infty dv
  I_\text{src}(v) 
  \left(\frac{\gamma-1}{\gamma}\right)
  \left( \frac{\gamma}{\gamma + u + v}\right)^\gamma 
  \times {}_2F_1 \left(\gamma/2,\frac{1+\gamma}{2},1,\frac{4uv}{(\gamma+u+v)^2}\right).
\end{align}
where $v=(r'/\sigma)^2/2$ and ${}_2F_1$ is the Gaussian hypergeometric
function.  This convolution formula reduces the expected photon
distribution to a single numerical integral.

@@

We estimate the error on the extension of a source by fixing
the position of the source and varying the extension until the
likelihood has fallen by \onehalf, corresponding to a $1\sigma$ error
\citep{Statistical_methods_book}.  Figure~\ref{four_plots_ic443}
demonstrates this method by showing the change in the log of the
likelihood when varying the extension of the SNR IC443.  The localization
error is calculated by fixing the extension and spectrum of the source
to their best fit value and then
fitting to the likelihood function a 2D Gaussian as a function
of position. This localization error algorithm is further described in
\cite{second_cat}.

@@

We simulated point-like sources with various spectral forms using
the LAT on-orbit simulation tool
\gtobssim\footnote{\gtobssim is distributed publicly by the FSSC.} and fit the sources
with \pointlike using both point-like
and extended source hypotheses.  These point-like sources were simulated with a power-law
spectral model with integrated fluxes above 100 \mev ranging from $3\times10^{-9}$ 
to $1\times10^{-5}$ \phflux and spectral
indices ranging from 1.5 to 3.  These values
were picked to represent typical parameters of LAT-detected
sources. The point-like sources were simulated on top of an isotropic
background with a power-law spectral model with
integrated flux above 100 \mev of $1.5\times10^{-5}$ \phflux
and spectral index 2.1.
This was
taken to be the same as the isotropic spectrum measured by EGRET
\citep{sreekumar_isotropic}.  The Monte Carlo simulation was performed
over a one-year observation period using a representative 
spacecraft orbit and livetime.
The reconstruction was performed
using the P7\_V6 Source class event selection and IRFs \citep{lat_on_orbit_psf}. For each 
significantly detected point-like source ($\ts\ge25$), we used \pointlike
to fit the source as an extended source and calculate \tsext.
\newtext{
This entire procedure was performed twice, once fitting in the 1 \gev
to 100 \gev energy range and once fitting in the 10 \gev to 100 \gev
energy range.

@@

Our study shows broad agreement between simulations and
Equation~\ref{ts_ext_distribution}. To the extent that there is
a discrepancy, the simulations tended to produce smaller than expected
values of \tsext which would make the formal significance conservative.
Considering the distribution in Figure~\ref{ts_ext_mc_1000} and
Figure~\ref{ts_ext_mc_10000}, the choice of a threshold \tsext set to 16
(correspond to a formal $4\sigma$ significance) is reasonable.
}

@@

We calculated the LAT flux threshold to detect that a spatially extended
source is extended. We define the detection threshold as the flux at
which the value of $\tsext$ averaged over many statistical realizations is
$\langle\tsext\rangle=16$ 
\newtext{
(corresponding to a formal $4\sigma$ significance)
for a source of a given flux and extension.
}

@@

After simultaneously fitting the two positions and two spectra,
we define \tsinc as twice the increase in the log of the likelihood
fitting the region with two point-like sources compared to fitting the
region with one point-like source:
\begin{equation}
  \tsinc=2\log(\likelihood_\text{2pts}/\likelihood_\text{ps}).
\end{equation} 
\newtext{
For the following analysis of LAT data, \tsinc was computed
by fitting the spectra of the two point-like sources in \gtlike using the best fit positions
of the sources found by \pointlike.
}

@@

We did this study first in the energy range from 1 \gev to 100 \gev by
simulating extended sources of flux $4\times10^{-9}$ \phflux integrated
from 1 \gev to 100 \gev and a power-law spectral model with
spectral index 2.  This flux was picked to be representative of the
new extended sources that were discovered in the following analysis
when looking in the 1 \gev to 100 \gev energy range
(see Section~\ref{new_ext_srcs_section}).
We simulated these sources using a uniform surface brightness profile
with extensions varying
up to $1\degree$.  
\newtext{
Figure~\ref{confusion_extended_plot}a shows the distribution
of \tsext and \tsinc and 
Figure~\ref{confusion_extended_plot}c shows 
the distribution of $\tsext-\tsinc$ as a
function of the simulated extension of the source
for 200 statistically independent simulations.

@@

Several interesting conclusions can be made from this study. The overall
behavior of these plots is what one would expect. $\tsext-\tsinc$
is mostly positive when fitting the simulated extended sources in
Figure~\ref{confusion_extended_plot} whereas $\tsext-\tsinc$ is
mostly negative when fitting the two simulated point-like sources in
Figure~\ref{confusion_2pts_plot}.  
Furthermore, $\tsext-\tsinc$ becomes more positive when the extended
soruce becomes larger and more negative when the two point-like sources
are farther apart.
From this, we conclude that the $\tsext>\tsinc$ test is
powerful at discarding cases in which the true emission comes from
two point-like sources without frequently rejecting real extended sources.

The other interesting feature in Figure~\ref{confusion_extended_plot}a
and \ref{confusion_extended_plot}b is that for simulated extended
sources with typical sizes ($\sigma\sim0\fdg5$), one can often obtain
almost as large an increase in likelihood fitting the source as two
point-like sources ($\tsinc\sim\tsext$).  This is because although the
two point-like sources represent an incorrect spatial model, the second
source has four additional degrees of freedom (two spatial and two
spectral parameters) and can therefore easy model much of the extended
source and statistical fluctuations in the data.  This effect is most
pronounced when using photons with energies between 1 \gev and 100 \gev
where the PSF is broader.
}

@@

For all following LAT analysis, we used the same two-year dataset
that was used in 2FGL spanning from 2008 August 4 to 2010 August 1. We
applied the same acceptance cuts and we used the same P7\_V6 Source class
event selection and IRFs \citep{lat_on_orbit_psf}.  
When analyzing sources in \pointlike, we used a circular $10\degree$ region of
interest (ROI) centered on our source and eight energy bins per
logarithmic decade in energy.
When refitting the region in \gtlike using the best fit spatial and
spectral models from \pointlike, we used the `binned likelihood' mode of
\gtlike on a $14\degree\times14\degree$ ROI with a pixel size of 0\fdg03.

Unless explicitly
mentioned, we used the same background model as 2FGL to represent the
Galactic diffuse, isotropic, and Earth limb emission.  To account for
possible residuals in the diffuse emission model, the galactic emission
was scaled by a power-law and the normalization
of the isotropic component
was left free.  
Unless explicitly mentioned,
we used all 2FGL sources within $15\degree$ of our source as our list
of background sources and we refit the spectral parameters of all sources
within $2\degree$ of the source.

@@

\newtext{
Of the 885 clean AGN, we selected the 733 of these 2FGL sources which
were significantly detected above 1 \gev and fit each of them for
extension.  The cumulative density of \tsext for these AGN is compared
to the $\chi^2_1/2$ distribution of Equation~\ref{ts_ext_distribution}
and is shown in Figure~\ref{agn_ts_ext}.  The \tsext distribution
for the AGN shows reasonable agreement with the theoretical
distribution and no AGN was found to be significantly extended
($\tsext>16$).  The observed discrepancy from the theoretical
distribution is likely due to small systematics
in our model of the LAT PSF and the Galactic diffuse emission (see
Section~\ref{systematic_errors_on_extension}).  The discrepancy could
also in a few cases be due to confusion with a nearby undetected source.
The overall agreement with the expected distribution demonstrates that
we can use \tsext as a measure of the statistical significance of the
detection of the extension of a source.
}

@@

Two nearby satellite galaxies of the Milky Way the Large Magellanic Cloud (LMC)
and the Small Magellanic
Cloud (SMC) were included in 2FGL as spatially extended sources \citep{lmc,smc}.  They were significantly
detected using photons with energies between
1 \gev and 100 \gev. Our
fit extension is comparable to the published result, but we note that
the previous LAT Collaboration publication on LMC used a more complicated two 2D Gaussian surface
brightness profile when fitting it \citep{lmc}.

% 3 PWN
Three PWNe, MSH\,15$-$52, Vela X, and HESS\,J1825$-$137, were
identified in 2FGL \citep{msh1552,velax,fermi_hess_j1825}.  
HESS\,J1825$-$137 was significantly detected using photons
with energies between 10 \gev and 100 \gev.
To improve our analysis of this source, removed
from our background model 
2FGL\,J1823.1$-$1338c which was included in 2FGL to
represent part of the extended emission.
To avoid confusion with the nearby bright pulsar PSR\,J1509$-$5850, MSH\,15$-$52 had
to be
analyzed at high energies.  Using photons with energies above 10 \gev,
we fit the extension of MSH\,15$-$52 to be consistent with the published
size but with \tsext=6.6.

@@

We developed two criteria for estimating systematic errors
on the extension of the sources.
First, we estimated a systematic error due to
uncertainty in our knowledge of the LAT PSF.  
Before launch, the LAT PSF
was determined by detector simulations which were verified in accelerator
beam tests \citep{atwood_LAT_mission}. However, in-flight data revealed
a discrepancy above a few \gev in the PSF compared to the angular
distribution of photons from bright AGN \citep{lat_on_orbit_psf}.
Subsequently, the PSF was fit empirically to bright AGN and this
empirical parameterization is used in the P7\_V6 IRFs.  To account for
our uncertainty in our knowledge of the PSF, we refit our extended source
candidates using the pre-flight Monte Carlo representation of the PSF
and consider the difference in extension found using the two PSFs as a
systematic error on the extension of a source.  The same approach was used
in \cite{ic443}.  At high energies, we believe that our parameterization
of the PSF from bright AGN is substantially better than the Monte Carlo
representation of the PSF so this systematic error is conservative.

% Alternate Diffuse

@@

It is not expected that this diffuse model is superior to the standard
LAT model obtained through an all-sky fit.  However, adding degrees of
freedom to the background model can remove likely spurious sources that
correlate with features in the Galactic diffuse emission.  Therefore,
this tests systematics that may be due to imperfect modeling of the
diffuse emission in the region. Nevertheless, this alternative diffuse
model does not test all systematics related to the diffuse emission model.
In particular, because the alternate model uses the same underlying gas
maps it is unable to assess systematics due to mischaracterization of local
diffuse structure due to lack of resolution of the underlying maps.

@@

\newtext{
We also performed a second search for extended sources using only
photons with energies between 10 \gev and 100 \gev.  Although this
approach tests the same sources, it is complementary because the galactic
diffuse emission is even less dominant above 10 \gev and because source
confusion is less of an issue.  A similar procedure was used to detect
the spatial extension of HESS\,j1825$-$137 and MSH\,15$-$52 with the
LAT \citep{msh1552,fermi_hess_j1825}.
}

@@

We found by visual inspection that in many
cases our analysis was strongly influenced by large-scale residuals in the
diffuse emission and hence the extension measure was unreliable.  This was
especially true in our analysis of sources in the 1 \gev to 100 \gev
energy range.  An example of such a case is 2FGL\,J1856.2+0450c analyzed
in the 1 \gev to 100 \gev energy range. Figure~\ref{example_bad_fit}
shows a diffuse-emission-subtracted smoothed counts map for
this source with the best
fit extension of the source overlaid. There appears to be large-scale
residual in the diffuse emission in this region along the Galactic plane.
As a result, 2FGL\,J1856.2+0450c is fit to an extension of $\sim2\degree$
and the result is statistically significant with \tsext=45.4. However,
by looking at the residuals it is clear that this complicated region is
not well modeled. We manually discard sources like this.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml
% (2FGL J1856.2+0450c = P72Y3047 for)

Because of these systematic issues, we only selected extended source
candidates in regions that did not appear dominated by these issues and
where there was a multiwavelength
counterpart. Because of these systematic issues, this search can not be
expected to be complete and it is likely that there are other spatially
extended sources that this method missed.

>>>>

\begin{abstract}
Spatial extension is an important characteristic for correctly
associating $\gamma$-ray-emitting sources with their counterparts at other wavelengths
and for obtaining an unbiased model of their spectra.  We present a new
method for quantifying the spatial extension of sources detected by the Large
Area Telescope (LAT), the primary science instrument on the {\em \fermi
Gamma-ray Space Telescope} (\fermi).  We perform a series of Monte Carlo
simulations to validate this tool and calculate the LAT threshold for
detecting the spatial extension of sources.  
We then test all sources
in the second \fermi-LAT catalog (2FGL) for extension. 
We report the detection of seven new spatially extended sources.
\end{abstract}

@@

A number of astrophysical source classes including supernova remnants
(SNRs), pulsar wind nebulae (PWNe), molecular clouds, normal galaxies,
and galaxy clusters are expected to be spatially resolvable at
\gev energies.  Additionally, dark matter satellites are also hypothesized to
be spatially extended.  The Large Area Telescope
(LAT) on the {\em \fermi Gamma-ray Space Telescope} (\fermi) has
detected seven SNRs which are significantly extended
at \gev energies: 
W51C, W30, 
IC443, W28, W44, RX\,J1713.7$-$3946,
and the Cygnus Loop
\citep{w51c,w30_lat,ic443,w28,w44,rx_j1713_lat,cygnus_loop_lat}. In addition, 
three extended PWN have been detected: MSH\,15$-$52, Vela~X, and
HESS\,J1825$-$137 \citep{msh1552,velax,fermi_hess_j1825}. Two
nearby galaxies, the Large and Small Magellanic Clouds, 
and the lobes of one
radio galaxy, Centaurus A, were spatially resolved at \gev energies
\citep{lmc,smc,cen_a_lat}.  A number of additional sources detected
at \gev energies are positionally coincident with sources that exhibit
large enough extension at other wavelengths to be spatially
resolvable by the LAT at \gev energies.

@@

A new maximum-likelihood analysis tool has been developed to address the
unique requirements for studying spatially extended sources with the LAT.
It works by maximizing the Poisson likelihood to find the observing counts
given a parametrized spatial and spectral model of the sky.  
Space is
divided into cubes of position and energy using the HEALPix representation of
the sky \citep{healpix} and the likelihood is maximized over all bins in
a region.
The extension of a source can be modeled by a geometric shape
(e.g. a disk or a two-dimensional Gaussian) and the position, extension,
and spectrum of the source can be simultaneously fit.

This type of analysis is not feasible using the standard LAT likelihood
analysis tool \gtlike\footnote{\gtlike is distributed publicly by the
FSSC.} because it can only fit the spectral parameters of the model
unless a more sophisticated iterative procedure is used.  
We note that \gtlike has been used in the
past in several studies of source extension in the LAT Collaboration
\citep{lmc,smc,w28,w51c}.  In these studies, a profile based upon a
set of \gtlike maximum likelihood fits at fixed extensions was used
to build a profile of the likelihood as a function of extension.
The \gtlike likelihood profile approach has been shown to correctly
reproduce the extension of simulated extended sources assuming that the
true position is known \citep{francesco_2011}.  But it is not optimal
because the position, extension, and spectrum of the source must be
simultaneously fit to find the best fit parameters and to maximize the
statistical significance of the detection.  Furthermore, because the \gtlike
approach is computationally intensive, no large-scale Monte Carlo
simulations have been run to calculate its false detection rate.

@@

In \pointlike, one can fit the position and extension 
of a source under the assumption that the source model
can be factorized:
$M(l,b,E)=S(l,b)\times X(E)$, where $S(l,b)$ is the spatial distribution
and $X(E)$ is the spectral distribution.  To fit an extended source,
\pointlike convolves the extended source shape with the PSF (as a function
of energy) and uses the \minuit library \citep{minuit_documentation}
to maximize the likelihood by simultaneously varying the position,
extension, and spectrum of the source.  As will be described in
Section~\ref{monte_carlo_validation}, simultaneously fitting the
position, extension, and spectrum is important to maximize
the statistical significance of the detection of the extension of a source.
To avoid projection effects, the longitude and latitude 
of the source are
not directly fit but instead the displacement of the source in a rotated
reference frame.

@@

For extended sources with an assumed radially-symmetric shape,
we optimized the calculation by performing one
of the integrals analytically.
The expected photon 
distribution can be written as
\begin{equation}
  \text{PDF}(\vec r) = \int  \text{PSF}(|\vec r - \vec r'|)I_\text{src}(\vec r') r' dr' d\phi'
\end{equation}
where $\vec r$ represents the position in the sky and
$I_\text{src}(\vec r)$ is the spatial distribution of the
source.
The PSF of the LAT is currently parameterized 
in the Pass~7\_V6 (P7\_V6) Source Instrument
Response Function \citep[IRFs,][]{lat_on_orbit_psf} by a King function \citep{king_function}:
\begin{equation}
  \text{PSF}(r) = 
  \frac{1}{2\pi\sigma^2}
  \left(1-\frac{1}{\gamma}\right)
  \left(1+\frac{u}{\gamma}\right)^{-\gamma},
\end{equation}
where $u=(r/\sigma)^2/2$ and $\sigma$ and $\gamma$ are free parameters
\citep{matthew_kerr_thesis}.  For radially-symmetric extended sources,
the angular part of the integral can be evaluated analytically
\begin{align}
  \text{PDF}(u) & = \int_0^\infty r' dr'
  I_\text{src}(v) 
  \int_0^{2\pi} d\phi' 
  \text{PSF}(\sqrt{2\sigma^2(u+v-2\sqrt{uv}\cos(\phi-\phi'))})
  \\
  & = \int_0^\infty dv
  I_\text{src}(v) 
  \left(\frac{\gamma-1}{\gamma}\right)
  \left( \frac{\gamma}{\gamma + u + v}\right)^\gamma 
  \times {}_2F_1 \left(\gamma/2,\frac{1+\gamma}{2},1,\frac{4uv}{(\gamma+u+v)^2}\right),
\end{align}
where $v=(r'/\sigma)^2/2$ and ${}_2F_1$ is the Gaussian hypergeometric
function.  This convolution formula reduces the expected photon
distribution to a single numerical integral.

@@

We estimate the error on the extension of a source by fixing
the position of the source and varying the extension until the
likelihood has fallen by 1/2, corresponding to a $1\sigma$ error
\citep{Statistical_methods_book}.  Figure~\ref{four_plots_ic443}
demonstrates this method by showing the change in the log of the
likelihood when varying the extension of the SNR IC443.  The localization
error is calculated by fixing the extension and spectrum of the source
to their best fit value and then
fitting to the likelihood function a 2D Gaussian as a function
of position. This localization error algorithm is further described in
\cite{second_cat}.

@@

We simulated point-like sources with various spectral forms using
the LAT on-orbit simulation tool
\gtobssim\footnote{\gtobssim is distributed publicly by the FSSC.} and fit the sources
with \pointlike using both point-like
and extended source hypotheses.  These point-like sources were simulated with a power-law
spectral model with integrated fluxes above 100 \mev ranging from $3\times10^{-9}$ 
to $1\times10^{-5}$ \phflux and spectral
indices ranging from 1.5 to 3.  These values
were picked to represent typical parameters of LAT-detected
sources. The point-like sources were simulated on top of an isotropic
background with a power-law spectral model with
integrated flux above 100 \mev of $1.5\times10^{-5}$ \phflux sr$^{-1}$
and spectral index 2.1.
This was
taken to be the same as the isotropic spectrum measured by EGRET
\citep{sreekumar_isotropic}.  The Monte Carlo simulation was performed
over a one-year observation period using a representative 
spacecraft orbit and livetime.
The reconstruction was performed
using the P7\_V6 Source class event selection and IRFs \citep{lat_on_orbit_psf}. For each 
significantly detected point-like source ($\ts\ge25$), we used \pointlike
to fit the source as an extended source and calculate \tsext.
This entire procedure was performed twice, once fitting in the 1 \gev
to 100 \gev energy range and once fitting in the 10 \gev to 100 \gev
energy range.

@@

Our study shows broad agreement between simulations and
Equation~\ref{ts_ext_distribution}. To the extent that there is
a discrepancy, the simulations tended to produce smaller than expected
values of \tsext which would make the formal significance conservative.
Considering the distribution in Figure~\ref{ts_ext_mc_1000} and
Figure~\ref{ts_ext_mc_10000}, the choice of a threshold \tsext set to 16
(correspond to a formal $4\sigma$ significance) is reasonable.

@@

We calculated the LAT flux threshold to detect that a spatially extended
source is extended. We define the detection threshold as the flux at
which the value of $\tsext$ averaged over many statistical realizations is
$\langle\tsext\rangle=16$ 
(corresponding to a formal $4\sigma$ significance)
for a source of a given extension.

@@

After simultaneously fitting the two positions and two spectra,
we define \tsinc as twice the increase in the log of the likelihood
fitting the region with two point-like sources compared to fitting the
region with one point-like source:
\begin{equation}
  \tsinc=2\log(\likelihood_\text{2pts}/\likelihood_\text{ps}).
\end{equation} 
For the following analysis of LAT data, \tsinc was computed
by fitting the spectra of the two point-like sources in \gtlike using the best fit positions
of the sources found by \pointlike.

@@

We did this study first in the energy range from 1 \gev to 100 \gev by
simulating extended sources of flux $4\times10^{-9}$ \phflux integrated
from 1 \gev to 100 \gev and a power-law spectral model with
spectral index 2.  This flux was picked to be representative of the
new extended sources that were discovered in the following analysis
when looking in the 1 \gev to 100 \gev energy range
(see Section~\ref{new_ext_srcs_section}).
We simulated these sources using a uniform surface brightness profile
with extensions varying
up to $1\degree$.  
Figure~\ref{confusion_extended_plot}a shows the distribution
of \tsext and \tsinc and 
Figure~\ref{confusion_extended_plot}c shows 
the distribution of $\tsext-\tsinc$ as a
function of the simulated extension of the source
for 200 statistically independent simulations.

@@

Several interesting conclusions can be made from this study. The overall
behavior of these plots is what one would expect. $\tsext-\tsinc$
is mostly positive when fitting the simulated extended sources in
Figure~\ref{confusion_extended_plot} whereas $\tsext-\tsinc$ is
mostly negative when fitting the two simulated point-like sources in
Figure~\ref{confusion_2pts_plot}.  
Furthermore, $\tsext-\tsinc$ becomes more positive when the extended
source becomes larger and more negative when the two point-like sources
are farther apart.
From this, we conclude that the $\tsext>\tsinc$ test is
powerful at discarding cases in which the true emission comes from
two point-like sources without frequently rejecting real extended sources.

The other interesting feature in Figure~\ref{confusion_extended_plot}a
and \ref{confusion_extended_plot}b is that for simulated extended
sources with typical sizes ($\sigma\sim0\fdg5$), one can often obtain
almost as large an increase in likelihood fitting the source as two
point-like sources ($\tsinc\sim\tsext$).  This is because although the
two point-like sources represent an incorrect spatial model, the second
source has four additional degrees of freedom (two spatial and two
spectral parameters) and can therefore easily model much of the extended
source and statistical fluctuations in the data.  This effect is most
pronounced when using photons with energies between 1 \gev and 100 \gev
where the PSF is broader.

@@

For all following LAT analyses, we used the same two-year dataset
that was used in 2FGL spanning from 2008 August 4 to 2010 August 1. We
applied the same acceptance cuts and we used the same P7\_V6 Source class
event selection and IRFs \citep{lat_on_orbit_psf}.  
When analyzing sources in \pointlike, we used a circular $10\degree$ region of
interest (ROI) centered on our source and eight energy bins per
logarithmic decade in energy.
When refitting the region in \gtlike using the best fit spatial and
spectral models from \pointlike, we used the `binned likelihood' mode of
\gtlike on a $14\degree\times14\degree$ ROI with a pixel size of 0\fdg03.

Unless explicitly
mentioned, we used the same background model as 2FGL to represent the
Galactic diffuse, isotropic, and Earth limb emission.  To compensate for
possible residuals in the diffuse emission model, the galactic emission
was scaled by a power-law and the normalization
of the isotropic component
was left free.  
Unless explicitly mentioned,
we used all 2FGL sources within $15\degree$ of our source as our list
of background sources and we refit the spectral parameters of all sources
within $2\degree$ of the source.

@@

Of the 885 clean AGN, we selected the 733 of these 2FGL sources which
were significantly detected above 1 \gev and fit each of them for
extension.  The cumulative density of \tsext for these AGN is compared
to the $\chi^2_1/2$ distribution of Equation~\ref{ts_ext_distribution}
in Figure~\ref{agn_ts_ext}.  The \tsext distribution
for the AGN shows reasonable agreement with the theoretical
distribution and no AGN was found to be significantly extended
($\tsext>16$).  The observed discrepancy from the theoretical
distribution is likely due to small systematics
in our model of the LAT PSF and the Galactic diffuse emission (see
Section~\ref{systematic_errors_on_extension}).  The discrepancy could
also in a few cases be due to confusion with a nearby undetected source.
The overall agreement with the expected distribution demonstrates that
we can use \tsext as a measure of the statistical significance of the
detection of the extension of a source.

@@

Two nearby satellite galaxies of the Milky Way the Large Magellanic Cloud (LMC)
and the Small Magellanic
Cloud (SMC) were included in 2FGL as spatially extended sources \citep{lmc,smc}.  
Their extension was significantly
detected using photons with energies between
1 \gev and 100 \gev. Our
fit extension is comparable to the published result, but we note that
the previous LAT Collaboration publication on LMC used a more complicated two 2D Gaussian surface
brightness profile when fitting it \citep{lmc}.

% 3 PWN
Three PWNe, MSH\,15$-$52, Vela X, and HESS\,J1825$-$137, were
identified in 2FGL \citep{msh1552,velax,fermi_hess_j1825}.  
HESS\,J1825$-$137 was significantly detected using photons
with energies between 10 \gev and 100 \gev.
To improve our analysis of this source, we removed
from our background model 
2FGL\,J1823.1$-$1338c which was included in 2FGL to
represent part of the extended emission.
To avoid confusion with the nearby bright pulsar PSR\,J1509$-$5850, MSH\,15$-$52 had
to be
analyzed at high energies.  Using photons with energies above 10 \gev,
we fit the extension of MSH\,15$-$52 to be consistent with the published
size but with \tsext=6.6.

@@

We developed two criteria for estimating systematic errors
on the extension of the sources.
First, we estimated a systematic error due to
uncertainty in our knowledge of the LAT PSF.  
Before launch, the LAT PSF
was determined by detector simulations which were verified in accelerator
beam tests \citep{atwood_LAT_mission}. However, in-flight data revealed
a discrepancy above a few \gev in the PSF compared to the angular
distribution of photons from bright AGN \citep{lat_on_orbit_psf}.
Subsequently, the PSF was fit empirically to bright AGN and this
empirical parameterization is used in the P7\_V6 IRFs.  To account for
the uncertainty in our knowledge of the PSF, we refit our extended source
candidates using the pre-flight Monte Carlo representation of the PSF
and consider the difference in extension found using the two PSFs as a
systematic error on the extension of a source.  The same approach was used
in \cite{ic443}.  At high energies, we believe that our parameterization
of the PSF from bright AGN is substantially better than the Monte Carlo
representation of the PSF so this systematic error is conservative.

% Alternative Diffuse Model

@@

It is not expected that this diffuse model is superior to the standard
LAT model obtained through an all-sky fit.  However, adding degrees of
freedom to the background model can remove likely spurious sources that
correlate with features in the Galactic diffuse emission.  Therefore,
this tests systematics that may be due to imperfect modeling of the
diffuse emission in the region. Nevertheless, this alternative diffuse
model does not test all systematics related to the diffuse emission model.
In particular, because the alternative model uses the same underlying gas
maps it is unable to assess systematics due to mischaracterization of local
diffuse structure due to lack of resolution of the underlying maps.

@@

We also performed a second search for extended sources using only
photons with energies between 10 \gev and 100 \gev.  Although this
approach tests the same sources, it is complementary because the galactic
diffuse emission is even less dominant above 10 \gev and because source
confusion is less of an issue.  A similar procedure was used to detect
the spatial extension of MSH\,15$-$52 and
HESS\,J1825$-$137 with the LAT \citep{msh1552,fermi_hess_j1825}.

@@

We found by visual inspection that in many
cases our analysis was strongly influenced by large-scale residuals in the
diffuse emission and hence the extension measure was unreliable.  This was
especially true in our analysis of sources in the 1 \gev to 100 \gev
energy range.  An example of such a case is 2FGL\,J1856.2+0450c analyzed
in the 1 \gev to 100 \gev energy range. Figure~\ref{example_bad_fit}
shows a diffuse-emission-subtracted smoothed counts map for
this source with the best
fit extension of the source overlaid. There appears to be large-scale
residuals in the diffuse emission in this region along the Galactic plane.
As a result, 2FGL\,J1856.2+0450c is fit to an extension of $\sim2\degree$
and the result is statistically significant with \tsext=45.4. However,
by looking at the residuals it is clear that this complicated region is
not well modeled. We manually discard sources like this.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml
% (2FGL J1856.2+0450c = P72Y3047 for)

We only selected extended source
candidates in regions that did not appear dominated by these issues and
where there was a multiwavelength
counterpart. Because of these systematic issues, this search can not be
expected to be complete and it is likely that there are other spatially
extended sources that this method missed.

For each candidate that was not biased by neighboring point-like
sources or by large-scale residuals in the diffuse emission model, we
improved the model of the region by deciding on a case by case basis which
background point-like sources should be kept.  We kept in our model of the sky
sources that we
believed represented physically distinct sources and we removed 
sources that we believed were included in 2FGL to compensate 
for residuals induced by not modeling the extension of the source.
Soft nearby point-like 2FGL sources that were not significant at higher energies
were frozen to the spectrum predicted by 2FGL.
When deciding
which background sources to keep and which to remove, we used as much as possible
multiwavelength information about possibly extended source counterparts
to help guide our choice. For each extended source presented in 
Section~\ref{new_ext_srcs_section}, we describe any modification from 2FGL
of the background model that were performed.
